---
title: "DraftDataScript"
author: "Heather Leonard, Abbie Sanders, and Rebecca Gordon" 
date: "11/18/2021"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
#installing gtsummary to help with creating summary tables
#install.packages("gtsummary")
```


```{r}
library(rio)
library(tidyverse)
library(here)
library(janitor)
library(gtsummary)
library(expss)
```
```{r}
synthdata <- read.csv(here ("data", "synthdata.csv")) %>%
  as_tibble()
```

## Notes for working on final project:
### Final project must use the following functions: pivot_longer(), select(), filter(), mutate(), pivot_wider(), group_by(), and summarize().

## Research question: Is internet overuse associated with sleep outcomes (total sleep, daytime sleepiness) in adolescents?

### Covariates: 

### Child age:
- dm_childage_bl
### Child sex:
- dm_sex_bl
### Welfare status:
- dm_lunch_bl 

### Steps:

1. select() and filter() will be used for narrowing down data to the variables we want
2. recode io variables will use mutate()
3. record age variable will use mutate()
4. create the composite score for io variables
5. in ggplot,  we can examine the relation between internet overuse and age by welfare status
6. summarize() by making a descriptives/summary table, possibly use pivots

### Make composite score for internet overuse, using variables io1_bl, io2_bl, io3_bl, and io4_bl. Levels include i don't know, never, rarely, sometimes, often, (1-5); new code is 0-4 scores with composite range of 0-16.

```{r abbie}
#recode internet overuse score and create a composite score 
synthdata <- synthdata %>%
  mutate(io1_RC = recode(io1_bl, "i don't know" = 0, "never" = 1, "often" = 2, "rarely" = 3, "sometimes" = 4))%>%
  mutate(io2_RC = recode(io2_bl, "i don't know" = 0, "never" = 1, "often" = 2, "rarely" = 3, "sometimes" = 4)) %>%
  mutate(io3_RC = recode(io3_bl, "i don't know" = 0, "never" = 1, "often" = 2, "rarely" = 3, "sometimes" = 4)) %>%
  mutate(io4_RC = recode(io4_bl, "i don't know" = 0, "never" = 1, "often" = 2, "rarely" = 3, "sometimes" = 4)) %>%
  mutate(io_comp = io1_RC + io2_RC + io3_RC + io4_RC)

#recode age
synthdata$age_RC <- as.numeric(str_sub(synthdata$dm_childage_bl, 1,2)) 

```


```{r pivotlonger-to-tidy-iodata}
synthdata %>%
  select(1:6, io1_RC:io4_RC)  %>%
  pivot_longer(
    cols = starts_with("io"),
    names_to = "variables",
    values_to = "answers"
  )
```


```{r}
#just playing around with pivotlonger here to see if we could come up with anything. it didn't work well. Need to come up with something.
synthdata %>% 
  pivot_longer(cols = grade, names_to = "variables", values_to = "answers")
```

```{r grouping-and-summarizing}
synthdata %>% 
  group_by(dm_sex_bl) %>% 
  summarise(mean_io = mean(io_comp))
```


```{r grouping-and-summarizing-using-pivot-wider}
synthdata %>% 
  group_by(dm_sex_bl) %>% 
  summarise(mean_io = mean(io_comp)) %>% 
  pivot_wider(names_from = dm_sex_bl, values_from = mean_io)
```


```{r summary-table-of-all-variables}
#this makes a giant table of all variables
synthdata %>% 
  tbl_summary()
```

```{r filter-characteristics-by-sex-and-show-pvalues}
#this will split out the characteristics by sex and add p value with second line. filtered out one gender answer of "other." 
synthdata %>% 
  select(grade, dm_childage_bl, dm_sex_bl, dm_ethnic_bl, dm_lunch_bl, dm_race_bl, sl_sleepy_bl, sl_avgsleepcat_bl, io_comp) %>% 
  apply_labels(synthdata, grade = "Grade", dm_childage_bl = "Age", dm_sex_bl = "Sex", dm_ethnic_bl = "Ethnicity", dm_lunch_bl = "Free/reduced Lunch", dm_race_bl = "Race", sl_sleepy_bl = "Daytime sleepiness Average", sl_avgsleepcat_bl = "Weekly hours of sleep per night", io_comp = " Mean Internet Overuse") %>%
  tbl_summary(by = dm_sex_bl) %>%
  bold_labels() %>%
  add_p()
```

```{r select-specific-variables-for-summary-table-with-pvalues}
synthdata %>% 
  select(sl_avgsleepcat_bl, io_comp) %>% 
  tbl_summary(by = sl_avgsleepcat_bl) %>% 
  add_p()
```

```{r select-sdifferent-variables-for-summary-table-with-pvalues}
synthdata %>% 
  select(sl_bedmedia_bl, io_comp) %>% 
  tbl_summary(by = sl_bedmedia_bl) %>% 
  add_p()
```
```{r correlation-of-bedtime-media-use-and-internet-overuse}
#answer to our research question: positive relationship that is very very small
cor(synthdata$sl_bedmedia_bl, synthdata$io_comp, method = "pearson", use = "complete.obs")

```

```{r correlation-ofinternet-overuse-and-total-sleep-on-schoolnights}
#not really a correlation between internet overuse and total sleep on school nights
cor(synthdata$io_comp, synthdata$sl_sleepschool_bl, method = "pearson", use = "complete.obs")
```


```{r correlation-of-internet-overuse-and-daytime-sleepiness}
#positive correlation that is very small
cor(synthdata$io_comp, synthdata$sl_sleepy_bl, method = "pearson", use = "complete.obs")
```
```{r correlation-of-internet-overuse-and-welfare-status}
#positive correlation that is very small
cor(synthdata$io_comp, as.numeric(synthdata$dm_lunch_bl), method = "pearson", use = "complete.obs")
```

```{r, Rebecca-ggplot}

#Removes all "i don't know" values
synthdata <- synthdata %>% 
filter(synthdata$age_RC != "i don't know" & synthdata$dm_lunch_bl != "i don't know")

synthdata %>% 
  ggplot(aes(io_comp, sl_sleepy_bl)) +
  geom_jitter(aes(color = age_RC)) +
  geom_smooth(method = "lm") +
  facet_wrap(~dm_lunch_bl) +
  labs(title = 'Relation between internet overuse and Daytime sleepiness', subtitle = 'Scatter plot by welfare status', x = 'Internet Overuse score', y = 'Daytime sleepiness', color = 'Age')

synthdata %>% 
  ggplot(aes(io_comp, sl_avgsleepcat_bl)) +
  geom_jitter(aes(color = age_RC)) +
  labs(title = 'Relation between internet overuse and Average Hours of sleep', x = 'Internet Overuse score', y = 'Hours of sleep', color = 'Age')

```
