---
title: "DraftDataScript"
author: "Heather Leonard, Abbie Sanders, and Rebecca Gordon" 
date: "11/18/2021"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
#installing gtsummary to help with creating summary tables
#install.packages("gtsummary")
```


```{r}
library(rio)
library(tidyverse)
library(here)
library(janitor)
library(gtsummary)

```
```{r}
synthdata <- read.csv(here ("data", "synthdata.csv")) %>%
  as_tibble()
```

Notes for working on final project:
Final project must use the following functions: pivot_longer(), select(), filter(), mutate(), pivot_wider(), group_by(), and summarize().

Research question: Is internet overuse associated with sleep outcomes (total sleep, daytime sleepiness) in adolescents?

possible covariates: dm_childage_bl, dm_sex_bl, dm_lunch_bl 

1. select() and filter() will be used for narrowing down data to the variables we want
2. recode io variables will use mutate()
3. record age variable will use mutate()
4. create the composite score for io variables
5. in ggplot we can use group_by()
6. summarize() by making a descriptives/summary table, possibly use pivots

Make composite score for internet overuse, using variables io1_bl, io2_bl, io3_bl, and io4_bl. Levels include i don't know, never, rarely, sometimes, often, (1-5); new code is 0-4 scores with composite range of 0-16.

```{r abbie}
#recode internet overuse score and create a composite score 
synthdata <- synthdata %>%
  mutate(io1_RC = recode(io1_bl, "i don't know" = 0, "never" = 1, "often" = 2, "rarely" = 3, "sometimes" = 4))%>%
  mutate(io2_RC = recode(io2_bl, "i don't know" = 0, "never" = 1, "often" = 2, "rarely" = 3, "sometimes" = 4)) %>%
  mutate(io3_RC = recode(io3_bl, "i don't know" = 0, "never" = 1, "often" = 2, "rarely" = 3, "sometimes" = 4)) %>%
  mutate(io4_RC = recode(io4_bl, "i don't know" = 0, "never" = 1, "often" = 2, "rarely" = 3, "sometimes" = 4)) %>%
  mutate(io_comp = io1_RC + io2_RC + io3_RC + io4_RC)

# Great job at including this all in one full code (using pipes); it makes it easy to understand.

#recode age
synthdata$age_RC <- as.numeric(str_sub(synthdata$dm_childage_bl, 1,2)) 

```
<!-- Great use of "$" above. We did not do this in our script, we wrote everything out 'longhand'. We will probably re-write our code using this format to simplify and shorten code chunks
-->

<<<<<<< HEAD
```{r, Rebecca_ggplot}

#Removes all "i don't know" values
synthdata <- synthdata %>% 
filter(synthdata$age_RC != "i don't know" & synthdata$dm_lunch_bl != "i don't know")

synthdata %>% 
  ggplot(aes(io_comp, age_RC)) +
  geom_jitter(aes(color = age_RC)) +
  geom_smooth(method = "lm") +
  facet_wrap(~dm_lunch_bl) +
  labs(title = 'Relation between internet overuse and age', subtitle = 'Scatter plot by welfare status', x = 'Internet Overuse score', y = 'Student age', color = 'Age')
# I would consider changing the dimensions of the visualization to make a bit wider; while also expanding the range of x-axis values. I think it would make the smooth line more telling for the viewer.
```

```{r pivotlonger-to-tidy-iodata}
synthdata %>%
  select(1:6, io1_RC:io4_RC)  %>%
  pivot_longer(
    cols = starts_with("io"),
    names_to = "variables",
    values_to = "answers"
  )
# Did this do anything to the dataframe? Are you saving this as an object/new dataframe to use? Or is this just to show using pivot-longer?
```

<!-- If you are trying to find a way to use pivot_longer() . . . Maybe use it for the responses to the different responses for sleep. I do not know your data well (or what many of the variables represent) but it seems that could be an easy way to show sleep by different times during the week.
OR
Could use to combine the race/ethnicity variables into a string under a single variable?
-->

```{r}
#just playing around with pivotlonger here to see if we could come up with anything. it didn't work well. Need to come up with something.
synthdata %>% 
  pivot_longer(cols = grade, names_to = "variables", values_to = "answers")
```

```{r grouping-and-summarizing}
synthdata %>% 
  group_by(dm_sex_bl, dm_childage_bl) %>% 
  summarise(mean_io = mean(io_comp))
# I added dm_childage_bl to this table. Not that it is really needed, but it gives a bit more information regarding differences based upon age.
```


```{r grouping-and-summarizing-using-pivot-wider}
synthdata %>% 
  group_by(dm_sex_bl) %>% 
  summarise(mean_io = mean(io_comp)) %>% 
  pivot_wider(names_from = dm_sex_bl, values_from = mean_io)

# Not sure of this summary will be useful since it tells the same information as your previous code-chunk in a less descriptive way.
```


```{r summary-table-of-all-variables}
#this makes a giant table of all variables
synthdata %>% 
  tbl_summary()
# This is an awesome function that I am totally going to use in my team's project. I would suggest renaming all of the variable names to what they actually represent so there is no issue with interpretation.
```

```{r filter-characteristics-by-sex-and-show-pvalues}
#this will split out the characteristics by sex and add p value with second line. filtered out one gender answer of "other." 
synthdata %>% 
  filter(dm_sex_bl != "other") %>% 
  tbl_summary(by = dm_sex_bl) %>% 
  add_p()

# again... awesome summary table that includes a p-value. Very simple but tells the whole picture of the data. Well done. My same note as above regarding variable names.
```

```{r select-specific-variables-for-summary-table-with-pvalues}
synthdata %>% 
  select(sl_avgsleepcat_bl, io_comp) %>% 
  tbl_summary(by = sl_avgsleepcat_bl) %>% 
  add_p()
```

```{r select-sdifferent-variables-for-summary-table-with-pvalues}
synthdata %>% 
  select(sl_bedmedia_bl, io_comp) %>% 
  tbl_summary(by = sl_bedmedia_bl) %>% 
  add_p()
```
```{r correlation-of-bedtime-media-use-and-internet-overuse}
#answer to our research question: positive relationship that is very very small
cor(synthdata$sl_bedmedia_bl, synthdata$io_comp, method = "pearson", use = "complete.obs")

# Awesome way to get at your research question, though I still do not know what the variable names represent. My one overarching comment is to rename the variables to understandable names for an unaware audience.
```



